Ignition of Fuel Beds by Firebrands

### 2022-03-28 manuscript4.tex
Worked on methodology for the multi heater manuscript.

\section{Abstract}
    some text
\section{Introduction}
    Influence of local disturbances both heated and non heated
\section{Methodology}
    The probability of flaming ignition for fuel beds consisting of Douglas-fir particles was measured for various configurations of two embers in a wind tunnel. The wind tunnel was operated at either 0.5 \si{\meter\per\second} or 5.8 \si{\meter\per\second} for each test series. Figure~\ref{fig:multiHeaterApparatus} shows a representation of the wind tunnel, fuel bed, and the automated lowering device. Table~\ref{tab:multiHeaterConfig} shows the test matrix for each series of tests. For all tests the heaters were spaced five diameters apart. This distance was chosen as it corresponded to the approximate size of the recirculation zone of a heater in a wind of 5.8 \si{\meter\per\second} oriented perpendicular to the flow. The wind speed was measured with a TSI-IFA300 hot wire probe.

            \rowcolors{2}{gray!25}{white}
        \begin{table}[hpbt]
            \normalsize
            \caption{Test matrix}
            \centering
            \begin{tabular}{ccccr}
                \rowcolor{gray!50}
               Test Series & Orientation & Heater A & Heater B & U\textsubscript{bulk} (\si{\meter\per\second})\\
                \hline
                1   & Parallel      & Hot  & Cold & 0.5 \\
                2   & Parallel      & Hot  & Cold & 5.8 \\
                3   & Parallel      & Hot  & Hot  & 0.5 \\
                4   & Parallel      & Hot  & Hot  & 5.8 \\
                5   & Perpendicular & Hot  & Cold & 0.5 \\
                6   & Perpendicular & Hot  & Cold & 5.8 \\
                7   & Perpendicular & Cold & Hot  & 0.5 \\
                8   & Perpendicular & Cold & Hot  & 5.8 \\
                9   & Perpendicular & Hot  & Hot  & 0.5 \\
                10  & Perpendicular & Hot  & Hot  & 5.8 
            \end{tabular}
            \label{tab:multiHeaterConfig}
        \end{table}

    The cartridge heaters used had diameters of 6.4\si{\milli\meter} and lengths of 51\si{\milli\meter}. The heater sizes were chosen to represent large firebrands with a high potential to ignite a fuel bed in a wildfire~\cite{Manzello2007}. The heater temperatures were controlled using a PID temperature controller implemented in LabVIEW. Heater temperatures ranged from 250\si{\celsius} to 750\si{\celsius}. Heater temperatures were measured using a type-K thermocouple attached to the center of the heater opposite the fuel bed.
    Controlling the temperature of the heater provides an advantage over natural burning or smoldering embers and pre-heated particles by removing the temperature variability of the heat source and enabling real time data logging of the heat source temperature. Controlling the temperature of the ember also removes some complexity of calculating the energy imparted to the fuel bed. The heater was lowered onto the fuel bed with an automated lowering device. The lowering device included a load cell and a PID controller to maintain a force equivalent to a two 10\si{\gram} firebrands throughout the experiment. To minimize flow disruptions due to apparatus the heaters were each attached to two 4-40 threaded rods that extended approximately 100\si{\milli\meter} below the load cell.

@@

    The energy imparted to the fuel bed by the heater(s) was estimated by applying an energy balance to the heater. The calculated heat fluxes to the fuel bed ranged between 5\si{\kilo\watt\per\square\meter} and 50\si{\kilo\watt\per\square\meter} per heater. These values align with those reported for heat fluxes from embers to instrumented substrates~\cite{Tao2020, Hernandez2018}. The power delivered to the heaters were measured with CR9580-10 current sensors and ZMPT101B voltage sensors. Temperature distributions were created from infrared images of the the heater taken with a FLIR SC6700 camera. A black body calibration was performed to correlate photon counts from the camera to temperature and produce a temperature distribution. Temperature profiles were created in the along the axis of the cylindrical heaters. The circumferential temperature of the heater was considered uniform.

>>>>

\section{Abstract}
    some text
\section{Introduction}
    Influence of local disturbances both heated and non heated
\section{Methodology}
    The probability of flaming ignition for fuel beds consisting of Douglas-fir particles was measured for eight configurations of two embers in a wind tunnel. The wind tunnel was operated at either 0.5 \si{\meter\per\second} or 5.8 \si{\meter\per\second} for each test series. The wind speed was measured with a TSI-IFA300 hot wire probe. Figure~\ref{fig:multiHeaterApparatus} shows a representation of the wind tunnel, fuel bed, and the automated lowering device. Table~\ref{tab:multiHeaterConfig} shows the test matrix for each series of tests. For all of the tests the heaters were oriented perpendicular to the flow and the downstream heater was heated. For tests where both heaters were heated the temperatures of both heaters were maintained at the same temperature.

    The combinations of heater spacing and hot or ambient upstream heater were chosen to represent different levels of fluid and thermal interactions between the heaters and the fuel bed. The heater spacing of five diameters was chosen such that minimal interaction between the heaters occurs as previous CFD calculations indicate that the recirculation zone of the upstream heater is approximately five diameters when the heater is in a wind of 5.8 \si{\meter\per\second} and an orientation perpendicular to the flow. Preliminary thermal calculations also indicated that the pyrolysis fronts created by each heater are unlikely to interact within previously observed times to ignition. The heater spacing of one diameter was chosen for a high level of interaction between both the fluid disturbances and thermal fronts of each ember. The one diameter spacing places the downstream heater inside the recirculation zone of the upstream heater under wind and the preliminary heat transfer calculations indicated that the thermal fronts of each heater will merge before the anticipated time to ignition. The chosen heater orientations also represent potential scenarios that may be encountered in a wildfire. The configuration with two heaters both heated is representative of a multi ember attack where the one diameter spacing approximates a firebrand pile and the five diameter spacing approximates two embers falling in close proximity but not within each others region of influence. The configuration with only the downstream heater heated is representative of an ember falling near an object (e.g., twig, rock, or cone) that disturbs the flow and may act as a heat sink for the energy deposited to the ember with the one and five diameter spacing representing an ember falling both within and outside of the region of influence of the ember. The configuration where the upstream heater is not heated and five diameters is also considered a control for comparison to previous single heater ignition results.

            \rowcolors{2}{gray!25}{white}
        \begin{table}[hpbt]
            \normalsize
            \caption{Test matrix}
            \centering
            \begin{tabular}{ccccr}
                \rowcolor{gray!50}
               Test Series & Heater Spacing & Upstream Heater & U\textsubscript{bulk} (\si{\meter\per\second})\\
                \hline
                1   & 1 & Ambient & 0.5 \\
                2   & 1 & Ambient & 5.8 \\
                3   & 1 & Hot     & 0.5 \\
                4   & 1 & Hot     & 5.8 \\
                5   & 5 & Ambient & 0.5 \\
                6   & 5 & Ambient & 5.8 \\
                7   & 5 & Hot     & 0.5 \\
                8   & 5 & Hot     & 5.8 \\
            \end{tabular}
            \label{tab:multiHeaterConfig}
        \end{table}

    The cartridge heaters used had diameters of 6.4\si{\milli\meter} and lengths of 51\si{\milli\meter}. The heater sizes were chosen to represent large firebrands with a high potential to ignite a fuel bed in a wildfire~\cite{Manzello2007}. The heater temperatures were controlled using a PID temperature controller implemented in LabVIEW. Heater temperatures ranged from 250\si{\celsius} to 750\si{\celsius}. Heater temperatures were measured using a type-K thermocouple attached to the center of each heater opposite the fuel bed.
    Controlling the temperature of the heater provides an advantage over natural burning or smoldering embers and pre-heated particles by removing the temperature variability of the heat source and enabling real time data logging of the heat source temperature. Controlling the temperature of the ember also removes some complexity of calculating the energy imparted to the fuel bed. The heater was lowered onto the fuel bed with an automated lowering device. The lowering device included a load cell and a PID controller to maintain a force equivalent to a two 10\si{\gram} firebrands throughout the experiment. To minimize flow disruptions due to apparatus the heaters were each attached to two 4-40 threaded rods that extended approximately 100\si{\milli\meter} below the load cell.

@@
